---
title: "Stan model results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I have created a total of 10 Stan models with differing levels of complexity. These models are as follows:

1. Simple no threshold model. -no threshold and a fixed intercept.
2. Simple no threshold varying intercept model. -no threshold but with varying intercepts via fish individual
3. Unhinged fixed threshold and intercept. -The discontinuous piecewise linear regression
4. Unhinged fixed threshold, varying intercept.
5. Unhinged fixed intercept varying threshold.
6. Unhinged varying threshold and intercept.
7. Hinged fixed threshold and intercept. -The continuous piecewise linear regression
8. Hinged varying intercept fixed threshold.
9. Hinged varying threshold fixed intercept.
10. Hinged varying threshold and intercept.

The varying slope unhinged models
11. Unhinged fixed threshold + intercept (varying slopes).
12. Unhinged fixed threshold, varying intercept (varying slopes).
13. Unhinged fixed intercept varying threshold (varying slopes).
14. Unhinged varying threshold and intercept (varying slopes)


##1. Simple no threshold model##
Notation:
$y_{i} = \alpha + \beta x_{i} + \epsilon$ 

(Good)
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~30s
- Max. Treedepth exceeded=0
- low BFMI=0
- Divergences=0

#Pairs plot#
```{r}
pairs(fit, pars=c("alpha","beta","epsilon","lp__", "yhat[110]"))
```

#Extracted#
```{r}
fishdat1 %>% 
  ggplot() +
  geom_point(aes(x = (prev)^2, y = (oto_size)^2, colour=maturity), alpha=0.1) +
  geom_point(aes(x = (prev)^2, y = mean_oto_size), colour="gold2", alpha=0.25, size=0.25) +
  geom_ribbon(aes(x=(prev)^2, ymin = lower, ymax = upper), alpha = 0.25)+
  theme_classic()
```

##2. Simple no threshold varying intercept model##
Notation:
$y_{i} = \alpha_{j \left[{i}\right]} + \beta x_{i} + \epsilon$

Max treedepth=10
Adapt Delta=0.8
- run time= ~3min
- Max. Treedepth exceeded= 499
- low BFMI=0
- Divergences=0

Worked well with a Max treedepth of 12:
Max treedepth=12
Adapt Delta=0.8
- run time= ~10min
- Max. Treedepth exceeded= 920
- low BFMI=0
- Divergences=0


(Good enough)
Max treedepth=13
Adapt Delta=0.8
- run time= ~10min
- Max. Treedepth exceeded= 196
- low BFMI=0
- Divergences=0
```{r}
pairs(two_fit, pars=c("alpha[110]","beta","epsilon","lp__", "sigma_alpha", "yhat[110]"))
```

#Extracted#



##3. Unhinged fixed threshold + intercept##
Notation:
$y_{i} = \left(\alpha_{1} +  \beta x_{i}\right) \left(1 - \tau_{i}\right) +  \left(\alpha_{2} +  \beta x_{i}\right)\tau_{i} + \epsilon $


$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{i}\right)}}$

constraint on bp. upper limit is 4
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~42min
- Max. Treedepth exceeded= 3935 
- low BFMI=1
- Divergences= 0

with bp normal but no upper constraint
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~49min
- Max. Treedepth exceeded= 3986
- low BFMI=0
- Divergences= 0

with bp as uniform
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~18min
- Max. Treedepth exceeded= 1000 
- low BFMI=1
- Divergences= 2403

using the complicated 3 model
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~46min
- Max. Treedepth exceeded= 3253 
- low BFMI=1
- Divergences= 0

using the 'Another possible unhinged' model
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~7min
- Max. Treedepth exceeded= 3986
- low BFMI=0
- Divergences= 0

using the 'Another possible unhinged' model
Max treedepth= 15
Adapt Delta= 0.8
- run time= ~264min
- Max. Treedepth exceeded= 3797
- low BFMI=0
- Divergences= 0

Change to a continuous breakpoint and run
(Good)
Max treedepth= 11
Adapt Delta= 0.8
- run time= ~3min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0


#Pairs plot#
```{r}
pairs(three_fit, pars=c("alpha1", "alpha2", "eta","beta", "epsilon","lp__", "yhat[110]"))
```

#Extracted#



####4. Unhinged fixed threshold, varying intercept###
Notation:

$y_{i} = \left(\alpha_{1_{{j}\left[{i}\right]}} +  \beta x_{i}\right) \left(1 - \tau_{i}\right) +  \left(\alpha_{2_{{j}\left[{i}\right]}} +  \beta x_{i}\right)\tau_{i} + \epsilon $

$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{i}\right)}}$

Max treedepth= 10
Adapt Delta= 0.8
- run time= ~60min
- Max. Treedepth exceeded=3905 
- low BFMI=2
- Divergences= 95

Upper bound of 4 for the bp.
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~92min
- Max. Treedepth exceeded= 3961 
- low BFMI=3
- Divergences= 39



Max treedepth= 10
Adapt Delta= 0.85
- run time= ~10min
- Max. Treedepth exceeded= 0 
- low BFMI=4
- Divergences= 1813

Max treedepth= 10
Adapt Delta= 0.875
- run time= ~27min
- Max. Treedepth exceeded= 966
- low BFMI=4
- Divergences= 1688

Change to a continuous breakpoint and run
(Not Good as there is a split eta (neither of which are in the correct place))
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~5min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0

#Pairs plot#
```{r}
pairs(four_fit, pars=c("alpha[110]", "delta", "eta","beta", "epsilon","lp__","mu_alpha" ,"sigma_alpha", "yhat[110]"))
```

#Extracted#


####5. Unhinged fixed intercept varying threshold###
Notation:

$y_{i} = \left(\alpha_{1} +  \beta x_{i}\right) \left(1 - \tau_{i}\right) +  \left(\alpha_{2} +  \beta x_{i}\right)\tau_{i} + \epsilon $

$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{{j\left[{i}\right]}}\right)}}$

Max treedepth= 10
Adapt Delta= 0.8
- run time= ~77min
- Max. Treedepth exceeded= 4000 
- low BFMI=0
- Divergences= 0

Max treedepth= 15
Adapt Delta= 0.8
- run time= ~62min
- Max. Treedepth exceeded= 4000 
- low BFMI=0
- Divergences= 0

Change to a continuous breakpoint
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~64min
- Max. Treedepth exceeded= 2008
- low BFMI= 0
- Divergences= 0

(Good)
Max treedepth= 11
Adapt Delta= 0.8
- run time= ~46min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0
#Pairs plot#
```{r}
pairs(five_fit, pars=c("alpha1", "alpha2", "eta[110]", "mu_eta", "sigma_eta","beta", "epsilon","lp__", "yhat[110]"))
```

#Extracted#


####6. Unhinged varying threshold and intercept###
Notation:

$y_{i} = \left(\alpha_{1_{{j}\left[{i}\right]}} +  \beta x_{i}\right) \left(1 - \tau_{i}\right) +  \left(\alpha_{2_{{j}\left[{i}\right]}} +  \beta x_{i}\right)\tau_{i} + \epsilon $

$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{{j\left[{i}\right]}}\right)}}$

Max treedepth= 10
Adapt Delta= 0.8
- run time= ~64min
- Max. Treedepth exceeded= 4000 
- low BFMI=0
- Divergences= 0

Change to a continuous breakpoint and run
(Great)
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~41min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0

#Pairs plot#
```{r}
pairs(six_fit, pars=c("alpha1[110]", "alpha2[110]", "eta[110]", "mu_eta", "sigma_eta","beta", "epsilon","lp__", "sigma_alpha", "yhat[110]"))
```

#Extracted#



####7. Hinged fixed threshold and intercept###
Notation:
$y_{i} = \alpha + \beta_{1} x_{i} + \beta_{2} \left(x_{i} - \eta\right) \tau_{i} + \epsilon$



$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{i}\right)}}$

Max treedepth= 10
Adapt Delta= 0.8
- run time= ~5min
- Max. Treedepth exceeded= 0 
- low BFMI=0
- Divergences= 272

Constrain bp's upper limit to 4
(Good model)-but wrong eta value
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~8min
- Max. Treedepth exceeded= 0 
- low BFMI=0
- Divergences= 0

After cutting the data to above the wrong breakpoint being discovered
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~7min
- Max. Treedepth exceeded= 583
- low BFMI=0
- Divergences= 519

Max treedepth= 10
Adapt Delta= 0.85
- run time= ~6min
- Max. Treedepth exceeded= 0
- low BFMI=0
- Divergences= 684

Max treedepth= 10
Adapt Delta= 0.9
- run time= ~4min
- Max. Treedepth exceeded= 9
- low BFMI=0
- Divergences= 62

Specified that the minimum value of the breakpoint was 0.15
Max treedepth= 10
Adapt Delta= 0.9
- run time= ~4min
- Max. Treedepth exceeded= 5
- low BFMI=0
- Divergences= 88

Change to a continuous breakpoint and run
(Good model)
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~1min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0

#Pairs plot#
```{r}
pairs(seven_fit, pars=c("alpha", "eta","beta1", "beta2", "epsilon","lp__", "yhat[110]", "slope_after", "intercept_after"))
```

#Extracted#



####8. Hinged varying intercept fixed threshold###
Notation:
$y_{i} = \alpha_{j \left[{i}\right]} + \beta_{1} x_{i} + \beta_{2} \left(x_{i} - \eta\right) \tau_{i} + \epsilon$



$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{i}\right)}}$

Max treedepth= 10
Adapt Delta= 0.8
- run time= ~9min
- Max. Treedepth exceeded= 0 
- low BFMI=0
- Divergences= 1005

Contrain bp upper limit to 4
(Good model)
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~17min
- Max. Treedepth exceeded= 0 
- low BFMI=0
- Divergences= 0

Change to a continuous breakpoint and run
(Good)
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~6min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0


#Pairs plot#
```{r}
pairs(eight_fit, pars=c("alpha[110]", "eta","beta1", "beta2", "epsilon","lp__", "sigma_alpha", "yhat[110]", "slope_after", "intercept_after[110]"))
```

#Extracted#



####9. Hinged varying threshold fixed intercept###
Notation:
$y_{i} = \alpha + \beta_{1} x_{i} + \beta_{2} \left(x_{i} - \eta_{j \left[{i}\right]}\right) \tau_{i} + \epsilon$



$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{{j\left[{i}\right]}}\right)}}$

Max treedepth= 10
Adapt Delta= 0.8
- run time= ~14min
- Max. Treedepth exceeded= 0 
- low BFMI=1
- Divergences= 0

Constrain upper limit of bp at 4
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~16min
- Max. Treedepth exceeded= 0 
- low BFMI=1
- Divergences= 0

Constrain upper limit of mu_bp at 4
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~14min
- Max. Treedepth exceeded= 0 
- low BFMI=2
- Divergences= 0

Change to a continuous breakpoint and run
Max treedepth= 10
Adapt Delta= 0.8
- run time= 
- Max. Treedepth exceeded= 
- low BFMI= 
- Divergences= 

#Pairs plot#
```{r}
pairs(nine_fit, pars=c("alpha", "eta[110]", "mu_eta", "sigma_eta","beta1", "beta2", "epsilon","lp__", "yhat[110]","slope_after", "intercept_after[110]"))
```

#Extracted#




####10. Hinged varying threshold and intercept###
Notation:
$y_{i} = \alpha_{j\left[{i}\right]} + \beta_{1} x_{i} + \beta_{2} \left(x_{i} - \eta_{j \left[{i}\right]}\right) \tau_{i} + \epsilon$



$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{{j\left[{i}\right]}}\right)}}$

Max treedepth= 10
Adapt Delta= 0.8
- run time= ~23min
- Max. Treedepth exceeded= 292
- low BFMI=4
- Divergences= 0

(Good)
Max treedepth= 12
Adapt Delta= 0.8
- run time= ~23min
- Max. Treedepth exceeded= 0
- low BFMI=4
- Divergences= 0

Change to a continuous breakpoint and run
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~20min
- Max. Treedepth exceeded= 188 
- low BFMI= 4
- Divergences= 112


(Good)
Max treedepth= 11
Adapt Delta= 0.8
- run time= ~16min
- Max. Treedepth exceeded= 0
- low BFMI= 4
- Divergences= 0


#Pairs plot#
```{r}
pairs(ten_fit, pars=c("alpha[110]", "eta[110]", "mu_eta", "sigma_eta","beta1", "beta2", "epsilon","lp__", "sigma_alpha", "yhat[110]", "slope_after", "intercept_after[110]"))
```

#Extracted#

```{r}
for (i in seq_along(ten_post$lp__)) {
  segments(x0 = min(fishdat$prev), x1 = ten_post$bp[i], 
           y0 = ten_post$intercept[i] + ten_post$beta1* min(fishdat$prev), 
           y1 = ten_post$intercept[i] + ten_post$beta1 * ten_post$bp[i], 
           col = alpha(3, .1))
  segments(x1 = max(fishdat$prev), x0 = ten_post$bp[i], 
           y1 = ten_post$intercept[i] + 
             ten_post$beta2[i] * (max(fishdat$prev) - ten_post$bp[i]) + 
             ten_post$beta1[i] * max(fishdat$prev), 
           y0 = ten_post$intercept[i] + ten_post$slope_after * ten_post$bp[i], 
           col = alpha(2, .1))
}
```


####11. Unhinged fixed threshold + intercept (varying slopes)###
Notation:

$y_{i} = \left(\alpha_{1} +  \beta_1 x_{i}\right) \left(1 - \tau_{i}\right) +  \left(\alpha_{2} +  \beta_2 x_{i}\right)\tau_{i} + \epsilon $


$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{i}\right)}}$

Change to a continuous breakpoint
(Good)
Max treedepth= 11
Adapt Delta= 0.8
- run time= ~6min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0

#Pairs plot#
```{r}
pairs(eleven_fit, pars=c("alpha1", "alpha2", "beta1", "beta2", "epsilon","lp__", "eta"))
```

#Extracted#

####12. Unhinged fixed threshold, varying intercept (varying slopes)###
Notation:

$y_{i} = \left(\alpha_{1_{{j}\left[{i}\right]}} +  \beta_1 x_{i}\right) \left(1 - \tau_{i}\right) +  \left(\alpha_{2_{{j}\left[{i}\right]}} +  \beta_2 x_{i}\right)\tau_{i} + \epsilon $

$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{i}\right)}}$



Change to a continuous breakpoint
(Not good because of split in eta)
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~5min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0

#Pairs plot#
```{r}
pairs(twelve_fit, pars=c("alpha1[110]", "alpha2[110]", "eta","beta1", "beta2", "epsilon","lp__", "sigma_alpha", "yhat[110]"))

```

#Extracted#


####13. Unhinged fixed intercept varying threshold (varying slopes)###
Notation:

$y_{i} = \left(\alpha_{1} +  \beta_1 x_{i}\right) \left(1 - \tau_{i}\right) +  \left(\alpha_{2} +  \beta_2 x_{i}\right)\tau_{i} + \epsilon $

$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{{j\left[{i}\right]}}\right)}}$


Change to a continuous breakpoint
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~43min
- Max. Treedepth exceeded= 2000
- low BFMI= 0
- Divergences= 0


Max treedepth= 11
Adapt Delta= 0.8
- run time= ~64min
- Max. Treedepth exceeded= 0
- low BFMI= 0
- Divergences= 0


#Pairs plot#
```{r}
pairs(thirteen_fit, pars=c("alpha1", "alpha2", "eta[110]", "mu_eta", "sigma_eta","beta1", "beta2", "epsilon","lp__", "yhat[110]"))
```

#Extracted#



####14. Unhinged varying threshold and intercept (varying slopes)###
Notation:

$y_{i} = \left(\alpha_{1_{{j}\left[{i}\right]}} +  \beta_1 x_{i}\right) \left(1 - \tau_{i}\right) +  \left(\alpha_{2_{{j}\left[{i}\right]}} +  \beta_2 x_{i}\right)\tau_{i} + \epsilon $

$\tau_{i} = \frac{1}{1 + e^{-100\left(x_{i} - \eta_{{j\left[{i}\right]}}\right)}}$


Change to a continuous breakpoint
Max treedepth= 10
Adapt Delta= 0.8
- run time= ~58min
- Max. Treedepth exceeded= 929
- low BFMI= 0
- Divergences= 0

(Good)
Max treedepth= 11
Adapt Delta= 0.8
- run time= ~47min
- Max. Treedepth exceeded= 2
- low BFMI= 0
- Divergences= 0

#Pairs plot#
```{r}
pairs(fourteen_fit, pars=c("alpha1[110]", "alpha2[110]", "eta[110]", "mu_eta", "sigma_eta","beta1", "beta2", "epsilon","lp__", "sigma_alpha", "yhat[110]"))
```

#Extracted#

